Tropical Mathematics for the Lambda Calculus

### 2023-01-16 section3B.tex
Proposal commit

\subsubsection{Bounded duplications}

It is known {\color{red}(([reference??])} that the maps $X\mapsto \C M_{\leq n}(X)$ ($n\in\BB N$) lift to functors $!_n:\LREL\to\LREL$.
The sequence $(!_n)_{n\in\N}$ can be then turned into a $\N$-graded linear exponential comonad on (the SMC) $\LREL$, satisfying the adjunction:
$\LREL(Z\otimes !_{n}X,Y) \simeq \LREL(Z, !_{n}X\multimap Y)$.
Therefore, $\LREL$, together with $(!_n)_{n\in\N}$, is a model for $\BSTLC$. 
Concretely, a type %$A::= o \mid !_{n}A \multimap A$
{\color{red}(i tipi meglio dirli in Section II)} is in interpreted as $\model o=\B 1$ and 
$\model{!_{n}A\multimap B}:= \C M_{\leq n}(\model A) \times \model B$.

Bounded types are interesting because of the following proposition:
\begin{proposition}
For all bounded types $A,B$, the morphisms from $\model A$ to $\model B$ (in all parametric relational models) correspond to polynomials.
\end{proposition}
\begin{proof}
It suffices to check that $\model A$ is finite for all bounded types $A$. Indeed this implies that a morphism $t:\model A\to \model B$ is a finite matrix $t: \model A \times \model B \to \Lawv$.
Hence, its corresponding map $\widehat t:\Lawv^{\model A} \to \Lawv^{\model B}$ is a polynomial.
\end{proof}

For example (here $!_{n}(\Lawv^{X}):= \Lawv^{\C M_{\leq n}(X)}$):
\begin{itemize}
\item a map $f\in \LREL( !_{1}\Lawv, \Lawv)$ is of the form $f(x)=\min \{x+a,b\}$;
\item a map $f\in \LREL(!_{2}\Lawv, \Lawv)$ is a ``quadratic'' polynomial $f(x)=\min\{2x+a, x+b, c\}$.
\end{itemize}

- tropical polynomials and tropical roots ?? Meglio in Section II

\subsubsection{Unbounded duplications}

@@

@@

>>>>

%\subsubsection{Unbounded duplications}

@@

{\color{red} Point to Sec VI}

@@

It is known {\color{red}(([reference??] e dire meglio)} that the maps $X\mapsto \C M_{\leq n}(X)$ ($n\in\BB N$) lift to functors $!_n:\LREL\to\LREL$.
The sequence $(!_n)_{n\in\N}$ can be then turned into a $\N$-graded linear exponential comonad on (the SMC) $\LREL$, satisfying the adjunction:
$\LREL(Z\otimes !_{n}X,Y) \simeq \LREL(Z, !_{n}X\multimap Y)$.
Therefore, $\LREL$, together with $(!_n)_{n\in\N}$, is a model for $\BSTLC$. 
Concretely, a type %$A::= o \mid !_{n}A \multimap A$
{\color{red}(i tipi meglio dirli in Section II)} is in interpreted as $\model o=\B 1$ and 
$\model{!_{n}A\multimap B}:= \C M_{\leq n}(\model A) \times \model B$.

%Bounded types are interesting because of the following proposition:
%\begin{proposition}
%For all bounded types $A,B$, the morphisms from $\model A$ to $\model B$ (in all parametric relational models) correspond to polynomials.
%\end{proposition}
%\begin{proof}
%It suffices to check that $\model A$ is finite for all bounded types $A$. Indeed this implies that a morphism $t:\model A\to \model B$ is a finite matrix $t: \model A \times \model B \to \Lawv$.Hence, its corresponding map $\widehat t:\Lawv^{\model A} \to \Lawv^{\model B}$ is a polynomial.
%\end{proof}

%For example (here $!_{n}(\Lawv^{X}):= \Lawv^{\C M_{\leq n}(X)}$):
%\begin{itemize}
%\item a map $f\in \LREL( !_{1}\Lawv, \Lawv)$ is of the form $f(x)=\min \{x+a,b\}$;
%\item a map $f\in \LREL(!_{2}\Lawv, \Lawv)$ is a ``quadratic'' polynomial $f(x)=\min\{2x+a, x+b, c\}$.
%\end{itemize}
